Modelling complex population structure using $F$-statistics and Principal Component Analysis

### 2021-07-11 fstats_pca_lewontin.tex

\section{Theory}
In this section, I will give a very brief formal introduction to $F$-statistics and PCA. A more detailed technical treatise of PCA is given in e.g. \cite{jolliffe2013}, and a useful guide to interpretation is \cite{cavalli-sforza1994}. Readers unfamiliar with $F$-statistics may find \cite{patterson2012}, \cite{peter2016} or \cite{oteo-garcia2021} helpful.

@@

A second difference is that frequently, rare SNPs are weighted higher in PCA, whereas all SNPs are weighted the same for $F$-statistics \citep{patterson2006, patterson2012}. This is a difference of convention; since $F$-statistics are summed over SNPs with the same expectation, $F$-statistics could also be calculated using the same weighting. The close connection between the two approaches developed here suggest that for most analyses, users might want to be consistent and use the same weighting for both types of analyses.

>>>>

\section{Theory}
In this section, I will give a very brief formal introduction to $F$-statistics and PCA. A more detailed technical treatise of PCA is given by e.g. \cite{jolliffe2013} and \cite{pachter2014}, and a useful guide to interpretation is \cite{cavalli-sforza1994}. Readers unfamiliar with $F$-statistics may find \cite{patterson2012}, \cite{peter2016} or \cite{oteo-garcia2021} helpful.

@@

A second difference is that frequently, rare SNPs are weighted higher in PCA, whereas all SNPs are weighted the same for $F$-statistics \citep{patterson2006, patterson2012}. This is a difference of convention \citep{cavalli-sforza1975}; since $F$-statistics are summed over SNPs with the same expectation, $F$-statistics could also be calculated using the same weighting. The close connection between the two approaches developed here suggest that for most analyses, users might want to be consistent and use the same weighting for both types of analyses.
